Vagueness, Noise and Communication

### 2014-08-09 paper/03_simulations.tex
Description of methodology for experiments.

\paragraph{Voronoiness.}
This measure aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
Recalling the results by \cite{Jager2007}, the stable outcomes of a population of agents playing a similarity-maximization game using pure strategies and evolving according to the replicator dynamics are those where ``the sender strategy is consistent with the Voronoi tessallation that is induced by the image of the receiver strategy.''~(\cite{Jager2007}, p. 562).
\mytodo{JPC}{Explain further?}
Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive measure to characterize that in terms of degrees.
With that in mind, and given $\Acts = \States$, we define Voronoiness as follows.
Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
\begin{multline*}
  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state, \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then } 1 \textrm{ else } 0
\end{multline*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States} \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat, \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg, \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{ else } 0
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.
\mytodo{JPC}{Give some intuitions on these?}
The measures are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.

>>>>

\paragraph{Voronoiness.}
This measure aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
Recalling the results by \cite{Jager2007}, the stable outcomes of a population of agents playing a similarity-maximization game using pure strategies and evolving according to the replicator dynamics are those where ``the sender strategy is consistent with the Voronoi tessallation that is induced by the image of the receiver strategy.''~(\cite{Jager2007}, p. 562).
\mytodo{JPC}{Explain further?}
Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive measure to characterize that in terms of degrees.
With that in mind, and given that this measure only makes sense for $\Acts = \States$, we define Voronoiness as follows.
Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
\begin{multline*}
  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state, \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then } 1 \textrm{ else } 0
\end{multline*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States} \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat, \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg, \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{ else } 0
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.
\mytodo{JPC}{Give some intuitions on these?}
The measures are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.

\subsection{Methodology}
\newcommand{\impairment}{\alpha}
In our simulations we focus on a concrete setup in line with \cite{Correia2013}.
Namely, we define a similarity-maximization game with $\States = \Acts$, where $\States$ is a discrete set of equidistant points within the unit interval $[0,1]$, including the boundaries (thus we always use $n \geq 2$).
As similarity measure $s:\States \times \States \rightarrow [0,1]$, we use a negative exponential of the of the square distance, \emph{i.e.}~for $\state,\state^\prime \in \States$:
\begin{align*}
  s(\state,\state^\prime) =
    \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \state = \state^\prime \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \state \neq \state^\prime \\
    e^{-\frac{\lVert \state - \state^\prime \rVert ^2}{\impairment^2}} & \textrm{otherwise} \\
    \end{cases}
\end{align*}
where $\impairment$ is an \emph{impairment} factor.
Here we deviate from \cite{Correia2013} by working with impairment $\impairment$ rather than acuity $\theta$, since we are interested in exploring the case of having $0$ impairment, and acuity does not have a natural maximum as defined.
Confusion is directly related with similarity, \emph{i.e.}~the more similar two states are, the more likely an agent is to perceive one as the other.
For simplicity, we equate one with the other by defining $C_{ij} = s(\state_i,\state_j)$.

We run the noise-perturbed replicator dynamics with random initial strategies $\Sstrat_0$ and $\Rstrat_0$ until a convergence criterion is met, namely that the total absolute change in both sender and receiver strategies does not amount to more than $0.01$, \emph{i.e.}:
\begin{align*}
  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left| \Sstrat^\prime(\state,\messg) - \Sstrat(\state,\messg) \right| < 0.01
\end{align*}
and
\begin{align*}
  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left| \Rstrat^\prime(\messg,\state) - \Rstrat(\messg,\state) \right| < 0.01 \,.
\end{align*}

We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
For each parameter combination, we perform $100$ trials.
